Erosion of somatic tissue identity with loss of the X-linked intellectual disability factor KDM5C

### 2024-09-09 submission/manuscript.Rmd
fig 5 legend

\begin{figure}
  \centering 
  \includegraphics{../submission/compiled_figs/KDM5C_ChIPseq.pdf}
  \caption[KDM5C binds to a subset of germline gene promoters during early embryogenesis]{\textbf{KDM5C binds to a subset of germline gene promoters during early embryogenesis A.} ChIPseeker localization of KDM5C peaks at different genomic regions in EpiLCs (top) and hippocampal and cortex primary neuron cultures (PNCs, bottom). \textbf{B.} Overlap of genes with KDM5C bound to their promoters in EpiLCs (purple) and PNCs (blue). \textbf{C.} Gene ontology (GO) comparision of genes with KDM5C bound to their promoter in EpiLCs and PNCs. Genes were classified as either bound in EpiLCs only (EpiLC only), unique to PNCs (PNC only, no significant ontologies) or bound in both PNCs and EpiLCs (shared). \textbf{D.} Average KDM5C binding around the transcription start site (TSS) of all germline-enriched genes in EpiLCs (left) and PNCs (right). \textbf{E.} KDM5C binding at the promoters of RNA-seq germline DEGs. Genes were classified as either only dysregulated in EpiLCs (EpiLC only), genes dyresgulated in the hippocampus or amygdala but not EpiLCs (brain only), or genes dysregulated in both EpiLCs and the brain (common). (Legend continued on next page.)}
  \label{figurelabel}
\end{figure}
\addtocounter{figure}{-1}
\begin{figure} [t!]
  \caption[]{\textbf{KDM5C binds to a subset of germline gene promoters during early embryogenesis.} (Legend continued.) \textbf{F.} Example KDM5C ChIP-seq bigwigs of DEGs unqiue to EpiLCs. Although both are expressed in \textit{Kdm5c}-KO EpiLCs, KDM5C is only bound to the \textit{Dazl} promoter and not the \textit{Stra8} promoter in EpiLCs. \textbf{G.} Example KDM5C ChIP-seq bigwigs of RNA-seq DEGs common between the brain and EpiLCs. KDM5C is bound to the \textit{D1Pas1} promoter but not the \textit{XXX} promoter in EpiLCs. \textbf{H.} Example KDM5C ChIP-seq bigwigs of RNA-seq DEGs unique to the brain. KDM5C is bound to the \textit{XXX} promoter but not the \textit{Meig1} promoter}
\end{figure}

\newpage
\begin{figure}
  \centering
  \includegraphics{../submission/compiled_figs/KDM5C_Mechanism.pdf}
  \caption[KDM5C’s catalytic activity promotes long-term silencing of germline genes via DNA methylation]{\textbf{KDM5C’s catalytic activity promotes long-term silencing of germline genes via DNA methylation.} \textbf{A.} Left - Bigwigs of representative histone 3 lysine 4 trimethylation (H3K4me3) ChIP-seq peaks at two germline genes in the wild-type (WT) and \textit{Kdm5c}-KO (5CKO) adult amygdala. Right - Average H3K4me3 at the transcripiton start site (TSS) of all germline-enriched genes in wild-type (dark red) and \textit{Kdm5c}-KO (light red) amygdala. \textbf{B.} Left - Bigwigs of representative histone 3 lysine 4 dimethylation (H3K4me2) ChIP-seq peaks at representative germline genes in wild-type and \textit{Kdm5c}-KO EpiLCs. Right - Average H3K4me2 at the TSS of all germline-enriched genes in wild-type (dark purple) and \textit{Kdm5c}-KO (light purple) EpiLCs. \textbf{C.} Diagram of embryonic stem cell (ESC) to epiblast-like cell (EpiLC) differentiation protocol and collection time points for RNA and protein. \textbf{D.} Real time quantitative PCR (RT-qPCR) of \textit{Kdm5c} RNA expression, calculated in comparision to TBP expression (2\textsuperscript{-deltaCT}). * p < 0.05, ** p < 0.01, *** p < 0.001, Welch's t-test. \textbf{E.} KDM5C protein expression normalized to DAXX. Quantified intensity using ImageJ (artificial units - au). Right - representative lanes of Western blot for KDM5C and DAXX. * p < 0.05, ** p < 0.01, *** p < 0.001, Welch's t-test \textbf{F.} XXX {G.} XXX.}
  \label{figurelabel}
\end{figure}

>>>>

\begin{figure}
  \centering 
  \includegraphics{../submission/compiled_figs/KDM5C_ChIPseq.pdf}
  \caption[KDM5C binds to a subset of germline gene promoters during early embryogenesis]{\textbf{KDM5C binds to a subset of germline gene promoters during early embryogenesis A.} ChIPseeker localization of KDM5C peaks at different genomic regions in EpiLCs (top) and hippocampal and cortex primary neuron cultures (PNCs, bottom). \textbf{B.} Overlap of genes with KDM5C bound to their promoters (TSS &#177; 500) in EpiLCs (purple) and PNCs (blue). \textbf{C.} Gene ontology (GO) comparision of genes with KDM5C bound to their promoter in EpiLCs and PNCs. Genes were classified as either bound in EpiLCs only (EpiLC only), unique to PNCs (PNC only, no significant ontologies) or bound in both PNCs and EpiLCs (shared). \textbf{D.} Average KDM5C binding around the transcription start site (TSS) of all germline-enriched genes in EpiLCs (left) and PNCs (right). \textbf{E.} Eulerr of number of germline-enriched genes (green) with significant KDM5C binding at their promoter in EpiLCs (purple). \textbf{F.} Example KDM5C ChIP-seq bigwigs of KDM5C binding at the \textit{Dazl} TSS but not \textit{Stra8} in EpiLCs. {G.} HOMER motif analysis of all KDM5C-bound germline gene promoters, highlighting significant enrichment of MAX, E2F6, and E2F motifs. \textbf{H.} Number of all gene promoters bound or unbound by KDM5C with instances of the E2F or E-box consensus sequence. \textbf{I.} HOMER motif analysis of all KDM5C-unbound germline gene promoters, highlighting significant enrichment of RFX family transcription factor motifs. \textbf{J.} Expression of RNA-seq DEG \textit{Rfx2} in wild-type and \textit{Kdm5c}-KO EpiLCs. P-value of Welch's t-test, expression in transcripts per million (TPM).}
  \label{figurelabel}
\end{figure}

\begin{figure}
  \centering
  \includegraphics{../submission/compiled_figs/KDM5C_Mechanism.pdf}
  \caption[KDM5C’s catalytic activity promotes long-term silencing of germline genes via DNA methylation]{\textbf{KDM5C’s catalytic activity promotes long-term silencing of germline genes via DNA methylation.} \textbf{A.} Left - Bigwigs of representative histone 3 lysine 4 trimethylation (H3K4me3) ChIP-seq peaks at two germline genes in the wild-type (WT) and \textit{Kdm5c}-KO (5CKO) adult amygdala. Right - Average H3K4me3 at the transcripiton start site (TSS) of all germline-enriched genes in wild-type (dark red) and \textit{Kdm5c}-KO (light red) amygdala. \textbf{B.} Left - Bigwigs of representative histone 3 lysine 4 dimethylation (H3K4me2) ChIP-seq peaks at representative germline genes in wild-type and \textit{Kdm5c}-KO EpiLCs. Right - Average H3K4me2 at the TSS of all germline-enriched genes in wild-type (dark purple) and \textit{Kdm5c}-KO (light purple) EpiLCs. \textbf{C.} Diagram of embryonic stem cell (ESC) to epiblast-like cell (EpiLC) differentiation protocol and collection time points for RNA and protein. \textbf{D.} Real time quantitative PCR (RT-qPCR) of \textit{Kdm5c} RNA expression, calculated in comparision to TBP expression (2\textsuperscript{-deltaCT}). * p < 0.05, ** p < 0.01, *** p < 0.001, Welch's t-test. \textbf{E.} KDM5C protein expression normalized to DAXX. Quantified intensity using ImageJ (artificial units - au). Right - representative lanes of Western blot for KDM5C and DAXX. * p < 0.05, ** p < 0.01, *** p < 0.001, Welch's t-test \textbf{F.} XXX {G.} XXX.}
  \label{figurelabel}
\end{figure}
